### 2013-05-13 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
updating parallel mc and spn chapters

\subsection{Leakage Fractions for Symmetric Systems}
\label{sec:symmetric_leakage_fractions}
When solving problems where the linear operator is symmetric, a host
of analytic theories exist based on the eigenvalue spectrum of the
operator that characterize their behavior in the context of
deterministic linear solvers. Using past work, these theories are
adapted to the domain decomposed adjoint Neumann-Ulam method using the
one-speed, two-dimensional neutron diffusion equation. In this subsection
we use a one-speed two-dimensional neutron diffusion system as the
symmetric model problem. Using the linear system generated by the
discretization of the model problem, we use a spectral analysis to
generate analytic relations for the eigenvalues of the operator based
on system parameters. Using the eigenvalue spectra, we then build
relationships to characterize the transport of stochastic histories in
a decomposed domain and the fraction of histories that leak from a
domain and will therefore have to be communicated. Finally, we compare
these analytic results to numerical experiments conducted with the
model problem.

\subsubsection{Symmetric Model Problem}
\label{subsec:model_problem}
For our numerical experiments, we choose the one-speed,
two-dimensional neutron diffusion equation as a model problem
\citep{duderstadt_nuclear_1976}:
\begin{equation}
  -\boldsymbol{\nabla} \cdot D \boldsymbol{\nabla} \phi + \Sigma_a
  \phi = S\:,
  \label{eq:diffusion_eq}
\end{equation}
where $\phi$ is the neutron flux, $\Sigma_a$ is the absorption cross
subsection, and $S$ is the source of neutrons. In addition, $D$ is the
diffusion coefficient defined as:
\begin{equation}
  D = \frac{1}{3 ( \Sigma_t - \bar{\mu}\Sigma_s )}\:,
  \label{eq:diffusion_coeff}
\end{equation}
where $\Sigma_s$ is the scattering cross subsection, $\Sigma_t = \Sigma_a
+ \Sigma_s$ is the total cross subsection, and $\bar{\mu}$ is the cosine
of the average scattering angle. For simplicity, we will take
$\bar{\mu} = 0$ for our analysis giving $D=(3 \Sigma_t)^{-1}$. In
addition, to further simplify we will assume a homogeneous domain such
that the cross subsections remain constant throughout. Doing this permits
us to rewrite Eq~(\ref{eq:diffusion_eq}) as:
\begin{equation}
  -D \boldsymbol{\nabla}^2 \phi + \Sigma_a \phi = S\:.
  \label{eq:diffusion_eq_simple}
\end{equation}

We choose a finite difference scheme on a square Cartesian grid to
discretize the problem. For the Laplacian, we choose the 9-point
stencil shown in Figure~\ref{fig:stencil} over a grid of size $h$
\citep{leveque_finite_2007}:
\begin{multline}
  \nabla^2_9\phi = \frac{1}{6h^2}[4 \phi_{i-1,j} + 4 \phi_{i+1,j}
    + 4 \phi_{i,j-1} + 4 \phi_{i,j+1} + \phi_{i-1,j-1}\\ +
    \phi_{i-1,j+1} + \phi_{i+1,j-1} + \phi_{i+1,j+1} - 20
    \phi_{i,j}]\:.
  \label{eq:nine_point_stencil}
\end{multline}
\begin{figure}[t!]
  \begin{center}
    \scalebox{1.25}{\input{chapters/mc_theory/stencil.pdftex_t}}
  \end{center}
  \caption{\textbf{Nine-point Laplacian stencil.}}
  \label{fig:stencil}
\end{figure}
We then have the following linear system to solve:
\begin{multline}
  -\frac{1}{6h^2}[4 \phi_{i-1,j} + 4 \phi_{i+1,j} + 4
    \phi_{i,j-1} + 4 \phi_{i,j+1} + \phi_{i-1,j-1}\\ + \phi_{i-1,j+1}
    + \phi_{i+1,j-1} + \phi_{i+1,j+1} - 20 \phi_{i,j}] + \Sigma_a
  \phi_{i,j} = s_{i,j}\:,
  \label{eq:fd_system}
\end{multline}
and in operator form:
\begin{equation}
  \ve{D}\boldsymbol{\phi}=\ve{s}\:,
  \label{eq:operator_system}
\end{equation}
where $\ve{D}$ is the diffusion operator, $\ve{s}$ is the source in
vector form and $\boldsymbol{\phi}$ is the vector of unknown fluxes.

To close the system, a set of boundary conditions is required. In the
case of a non-reentrant current condition applied to all global
boundaries of the domain, we choose the formulation of Duderstadt by
assuming the flux is zero at some ghost point beyond the
grid. Consider for example the equations on the $i=0$ boundary of the
domain:
\begin{multline}
  -\frac{1}{6h^2}[4 \phi_{-1,j} + 4 \phi_{1,j} + 4 \phi_{0,j-1} +
    4 \phi_{0,j+1} + \phi_{-1,j-1}\\ + \phi_{-1,j+1} + \phi_{1,j-1} +
    \phi_{1,j+1} - 20 \phi_{0,j}] + \Sigma_a \phi_{0,j} = s_{0,j}\:.
  \label{eq:x_min_bnd}
\end{multline}
Here we note some terms where $i=-1$ and therefore are representative
of grid points beyond the boundary of the domain. We set the flux at
these points to be zero, giving a valid set of equations for the $i=0$
boundary:
\begin{multline}
  -\frac{1}{6h^2}[4 \phi_{1,j} + 4 \phi_{0,j-1} + 4 \phi_{0,j+1}
    \\ + \phi_{-1,j+1} + \phi_{1,j-1} + \phi_{1,j+1} - 20 \phi_{0,j}]
  + \Sigma_a \phi_{0,j} = s_{0,j}\:.
  \label{eq:x_min_bnd_2}
\end{multline}
We repeat this procedure for the other boundaries of the domain. For
reflecting boundary conditions, the net current across a boundary is
zero.

@@

The spectral radius of the iteration matrix is obtained by seeking its
largest eigenvalue. As with the diffusion operator, we can use the
same analysis techniques to find the eigenvalues for the iteration
matrix. We use a few simplifications by noting that if the Jacobi
preconditioned iteration matrix is $\ve{H} = \ve{I} -
\ve{M}^{-1}\ve{D}$, the we except all terms on the diagonal of the
iteration matrix to be zero such that we have the following stencil:
\begin{multline}
  \ve{H}\boldsymbol{\phi} = \frac{\alpha D}{6h^2}\Big[4 \phi_{i-1,j} +
    4 \phi_{i+1,j} + 4 \phi_{i,j-1} + 4 \phi_{i,j+1}
    +\\ \phi_{i-1,j-1} + \phi_{i-1,j+1} + \phi_{i+1,j-1} +
    \phi_{i+1,j+1}\Big]\:.
  \label{eq:iteration_stencil}
\end{multline}
Inserting the eigenfunctions defined by
Eq~(\ref{eq:eigenfunction_form}) we get:
\begin{multline}
  \lambda_{p,q}(\ve{H}) = \frac{\alpha D}{6h^2}\Big[4 e^{-2 \pi \imath p
      h} + 4 e^{2 \pi \imath p h} + 4 e^{-2 \pi \imath q h} + 4 e^{2
      \pi \imath q h} + e^{-2 \pi \imath p h} e^{-2 \pi \imath q h}
    \\ + e^{-2 \pi \imath p h} e^{2 \pi \imath q q} + e^{2 \pi \imath
      p h} e^{-2 \pi \imath q h} + e^{2 \pi \imath p h} e^{2 \pi
      \imath q h}\Big]\:,
  \label{eq:iteration_deriv}
\end{multline}
which simplifies to:
\begin{equation}
  \lambda_{p,q}(\ve{H}) = \frac{\alpha D}{6h^2}[ 8 \cos(\pi p h) + 8
    \cos(\pi q h) + 4 \cos(\pi p h) \cos(\pi q h)]\:,
  \label{eq:iteration_spectrum}
\end{equation}
giving the eigenvalue spectrum for the Jacobi preconditioned iteration
matrix. To find these maxium eigenvalue,
Eq~(\ref{eq:preconditioned_eigenvalues}) is plotted as a function of
$p$ with $p=q$ in Figure~\ref{fig:diffusion_spectrum} for various
values of $\Sigma_a$.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{chapters/mc_theory/diffusion_spectrum.png}
  \end{center}
  \caption{\textbf{Eigenvalue spectra for the diffusion equation.}}
  \label{fig:diffusion_spectrum}
\end{figure}
We find that the maximum eigenvalue exists when $p=q=0$, giving the
following for the spectral radius of the Jacobi preconditioned
iteration matrix:
\begin{equation}
  \rho(\ve{H}) = \frac{10 \alpha D}{3 h^2}\:.
  \label{eq:iteration_radius}
\end{equation}

@@

\subsubsection{Numerical Experiments}
\label{subsec:numerical_experiments}
To test the relationships developed by the spectral analysis, we form
two simple numerical experiments using the diffusion model problem:
one to measure the length of the random walks as a function of the
iteration matrix eigenvalues, and one to measure the domain leakage
fraction as a function of the iteration matrix eigenvalues and the
discretization properties. Before doing this, we verify our
computation of the spectral radius of the iteration matrix by
numerically computing the largest eigenvalue of the diffusion operator
using an iterative eigenvalue solver. For this verification, a $100
\times 100$ square grid with $h=0.01$, $h=0.1$, and $h=1.0$ and the
absorption cross varied from 0 to 100 while the scattering cross
subsection was fixed at unity. Figure~\ref{fig:measured_spec_rad} gives
the measured spectral radius of the iteration matrix and the computed
spectral radius for the preconditioned diffusion operator using
Eq~(\ref{eq:iteration_radius}) as function of the absorption to
scattering ratio $(\Sigma_a / \Sigma_s)$. Excellent agreement was
observed between the analytic and numerical results with all data
points computed within the tolerance of the iterative eigenvalue
solver.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/measured_spec_rad.png}
    \end{center}
    \caption{\textbf{Measured and analytic preconditioned diffusion
        operator spectral radius as a function of the absorption cross
        subsection to scattering cross subsection ratio.} \textit{Values of
        $h=0.01$, $h=0.1$, and $h=1.0$ were used. The red data was
        computed numerically by an eigensolver while the black dashed
        data was generated by Eq~(\ref{eq:iteration_radius}).}}
    \label{fig:measured_spec_rad}
  \end{spacing}
\end{figure}

\subsubsection{Random Walk Length}
\label{subsubsec:walk_length}
With the eigenvalue derivations verified, we can go about setting up
an experiment to measure the length of the random walks generated by
the adjoint Neumann-Ulam solver. To do this, we again use a $100
\times 100$ square grid with $h=0.1$ and the absorption cross varied
from 0 to 100 while the scattering cross subsection was fixed at
unity. Three weight cutoff values of \sn{1}{-2}, \sn{1}{-4}, and
\sn{1}{-8} were used with 10,000 histories generated by a point source
of strength 1 in the center of the domain. For each of the histories,
the number of transitions made was tallied to provide an effective
value of $k$ for each history. This value was then averaged over all
histories to get a measured value of $k$ for the particular
operator. Figure~\ref{fig:measured_length} presents these measurements
as well as the analytic result computed by Eq~(\ref{eq:analytic_k}) as
a function of the iteration matrix spectral radius,
$\rho(\ve{H})$. Figure~\ref{fig:measured_length_error} gives the
relative error between the predicted and observed results. We note
good qualitative agreement between the measured and analytic
results. However, we observe a larger relative error for both long and
short random walks.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length as a
        function of the iteration matrix spectral radius.} \textit{The
        weight cutoff was varied with \sn{1}{-2}, \sn{1}{-4}, and
        \sn{1}{-8}. The red data was computed numerically by an
        adjoint Neumann-Ulam implementation while the black dashed
        data was generated by Eq~(\ref{eq:analytic_k}).}}
    \label{fig:measured_length}
  \end{spacing}
\end{figure}
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/walk_length_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length relative
        error as a function of the iteration matrix spectral radius.}
      \textit{The weight cutoff was varied with \sn{1}{-2},
        \sn{1}{-4}, and \sn{1}{-8}. The relative error between the
        predicted and measured results is presented for each weight
        cutoff.}}
    \label{fig:measured_length_error}
  \end{spacing}
\end{figure}

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the center domain was computed and as the stage
progressed, the number of histories that exited that domain was
tallied with the ratio of the two numbers providing a numerical
measure for the leakage fraction. Figure~\ref{fig:measured_leakage}
gives the domain leakage measurements for the domain in the center of
the global grid as well as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5.25in,clip]{chapters/mc_theory/leakage_variation.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a
        function of the iteration matrix spectral radius.} \textit{To
        test the behavior with respect to domain size, $n_i=50$,
        $n_i=100$,and $n_i=200$ were used. The red data was computed
        numerically by a domain-decomposed adjoint Neumann-Ulam
        implementation, the black dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated
        by Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner
        rational approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute error between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/mc_theory/leakage_error.png}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute
        error as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size,
        $n_i=50$ (green), $n_i=100$ (blue), and $n_i=200$ (red) were
        used. The dashed lines represent the error using the Wigner
        rational approximation while the solid lines represent the
        error using the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these error results, the mean chord approximation is shown to
have a lower error for ill-conditioned systems as compared to the
Wigner approximation while the Wigner approximation produces less
error for more well-conditioned systems. We also note that for the
optically thick domains, the error is likely corresponded to that
observed in Figure~\ref{fig:measured_length} for the $k$ parameter
while the large relative error in $k$ for optically thin domains does
not affect the approximation significantly. In general, the mean chord
approximation is a better choice to estimate the leakage fraction in a
domain from the adjoint Neumann-Ulam method and except for a single
data point with $n_i=50$, the mean chord approximation yielded leakage
fractions within 0.05 of the measured results. As the domain becomes
more optically thick (with both increasing $n_i$ and decreasing
$\rho(\ve{H})$), the approximations are more accurate.

@@

\subsection{Transition Processing Kernel}
\label{subsec:transition_processing}
Processing the transition of a single stochastic history from one
state to another in the system is the fundamental kernel of the
transport procedure. In this kernel, an incoming history has a state
consisting of the current discrete state in the system in which it
resides and its current weight and is processed as given by
Algorithm~\ref{alg:transition_process}.
\begin{algorithm}[h!]
  \caption{Stochastic History Transition Processing Kernel}
  \label{alg:power_iteration}
  \begin{algorithmic}
    \State processTransition( history ):
    \State i = history.state
    \State j = sampleRowCDF( C[i] )
    \State history.state = j
    \State history.weight *= w[i,j]
    \EndWhile
  \end{algorithmic}
\end{algorithm}
The core component here is the sampling of the CDF for a particular
state generated by the probablities from the Neumann-Ulam
decomposition. Upon exiting, the kernel has modified the state of the
history with a new state and multiplied the weight with the
corresponding transition weight to form the product in
Eq~(\ref{eq:direct_permutation_weight}).

>>>>

\subsection{Leakage Fractions for Symmetric Systems}
\label{sec:symmetric_leakage_fractions}
When solving problems where the linear operator is symmetric, a host
of analytic theories exist based on the eigenvalue spectrum of the
operator that characterize their behavior in the context of
deterministic linear solvers. Using past work, these theories are
adapted to the domain decomposed adjoint Neumann-Ulam method using the
one-speed, two-dimensional neutron diffusion equation and spatial
discretization presented in \S~\ref{subsec:model_problem}. Using the
linear system generated by the discretization of the model problem, we
use a spectral analysis to generate analytic relations for the
eigenvalues of the operator based on system parameters. Using the
eigenvalue spectra, we then build relationships to characterize the
transport of stochastic histories in a decomposed domain and the
fraction of histories that leak from a domain and will therefore have
to be communicated. Finally, we compare these analytic results to
numerical experiments conducted with the model problem.

@@

The spectral radius of the iteration matrix is obtained by seeking its
largest eigenvalue. As with the diffusion operator, we can use the
same analysis techniques to find the eigenvalues for the iteration
matrix. We use a few simplifications by noting that if the Jacobi
preconditioned iteration matrix is $\ve{H} = \ve{I} -
\ve{M}^{-1}\ve{D}$, the we except all terms on the diagonal of the
iteration matrix to be zero such that we have the following stencil:
\begin{multline}
  \ve{H}\boldsymbol{\phi} = \frac{\alpha D}{6h^2}\Big[4 \phi_{i-1,j} +
    4 \phi_{i+1,j} + 4 \phi_{i,j-1} + 4 \phi_{i,j+1}
    +\\ \phi_{i-1,j-1} + \phi_{i-1,j+1} + \phi_{i+1,j-1} +
    \phi_{i+1,j+1}\Big]\:.
  \label{eq:iteration_stencil}
\end{multline}
Inserting the eigenfunctions defined by
Eq~(\ref{eq:eigenfunction_form}) we get:
\begin{multline}
  \lambda_{p,q}(\ve{H}) = \frac{\alpha D}{6h^2}\Big[4 e^{-2 \pi \imath p
      h} + 4 e^{2 \pi \imath p h} + 4 e^{-2 \pi \imath q h} + 4 e^{2
      \pi \imath q h} + e^{-2 \pi \imath p h} e^{-2 \pi \imath q h}
    \\ + e^{-2 \pi \imath p h} e^{2 \pi \imath q q} + e^{2 \pi \imath
      p h} e^{-2 \pi \imath q h} + e^{2 \pi \imath p h} e^{2 \pi
      \imath q h}\Big]\:,
  \label{eq:iteration_deriv}
\end{multline}
which simplifies to:
\begin{equation}
  \lambda_{p,q}(\ve{H}) = \frac{\alpha D}{6h^2}[ 8 \cos(\pi p h) + 8
    \cos(\pi q h) + 4 \cos(\pi p h) \cos(\pi q h)]\:,
  \label{eq:iteration_spectrum}
\end{equation}
giving the eigenvalue spectrum for the Jacobi preconditioned iteration
matrix. To find these maxium eigenvalue,
Eq~(\ref{eq:preconditioned_eigenvalues}) is plotted as a function of
$p$ with $p=q$ in Figure~\ref{fig:diffusion_spectrum} for various
values of $\Sigma_a$.
\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=5in,clip]{chapters/parallel_mc/diffusion_spectrum.png}
  \end{center}
  \caption{\textbf{Eigenvalue spectra for the diffusion equation.}}
  \label{fig:diffusion_spectrum}
\end{figure}
We find that the maximum eigenvalue exists when $p=q=0$, giving the
following for the spectral radius of the Jacobi preconditioned
iteration matrix:
\begin{equation}
  \rho(\ve{H}) = \frac{10 \alpha D}{3 h^2}\:.
  \label{eq:iteration_radius}
\end{equation}

@@

\subsubsection{Numerical Experiments}
\label{subsec:numerical_experiments}
To test the relationships developed by the spectral analysis, we form
two simple numerical experiments using the diffusion model problem:
one to measure the length of the random walks as a function of the
iteration matrix eigenvalues, and one to measure the domain leakage
fraction as a function of the iteration matrix eigenvalues and the
discretization properties. Before doing this, we verify our
computation of the spectral radius of the iteration matrix by
numerically computing the largest eigenvalue of the diffusion operator
using an iterative eigenvalue solver. For this verification, a $100
\times 100$ square grid with $h=0.01$, $h=0.1$, and $h=1.0$ and the
absorption cross varied from 0 to 100 while the scattering cross
subsection was fixed at unity. Figure~\ref{fig:measured_spec_rad} gives
the measured spectral radius of the iteration matrix and the computed
spectral radius for the preconditioned diffusion operator using
Eq~(\ref{eq:iteration_radius}) as function of the absorption to
scattering ratio $(\Sigma_a / \Sigma_s)$. Excellent agreement was
observed between the analytic and numerical results with all data
points computed within the tolerance of the iterative eigenvalue
solver.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=5in,clip]{chapters/parallel_mc/measured_spec_rad.png}
    \end{center}
    \caption{\textbf{Measured and analytic preconditioned diffusion
        operator spectral radius as a function of the absorption cross
        subsection to scattering cross subsection ratio.} \textit{Values of
        $h=0.01$, $h=0.1$, and $h=1.0$ were used. The red data was
        computed numerically by an eigensolver while the black dashed
        data was generated by Eq~(\ref{eq:iteration_radius}).}}
    \label{fig:measured_spec_rad}
  \end{spacing}
\end{figure}

\subsubsection{Random Walk Length}
\label{subsubsec:walk_length}
With the eigenvalue derivations verified, we can go about setting up
an experiment to measure the length of the random walks generated by
the adjoint Neumann-Ulam solver. To do this, we again use a $100
\times 100$ square grid with $h=0.1$ and the absorption cross varied
from 0 to 100 while the scattering cross subsection was fixed at
unity. Three weight cutoff values of \sn{1}{-2}, \sn{1}{-4}, and
\sn{1}{-8} were used with 10,000 histories generated by a point source
of strength 1 in the center of the domain. For each of the histories,
the number of transitions made was tallied to provide an effective
value of $k$ for each history. This value was then averaged over all
histories to get a measured value of $k$ for the particular
operator. On the left, Figure~\ref{fig:measured_length} presents these
measurements as well as the analytic result computed by
Eq~(\ref{eq:analytic_k}) as a function of the iteration matrix
spectral radius, $\rho(\ve{H})$. On the right,
Figure~\ref{fig:measured_length} gives the relative difference between the
predicted and observed results. We note good qualitative agreement
between the measured and analytic results. However, we observe a
larger relative difference for both long and short random walks.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=6.0in,clip]{chapters/parallel_mc/measured_length_2.pdf}
    \end{center}
    \caption{\textbf{Measured and analytic random walk length as a
        function of the iteration matrix spectral radius.} \textit{The
        weight cutoff was varied with \sn{1}{-2}, \sn{1}{-4}, and
        \sn{1}{-8}. In the left plot, the red data was computed
        numerically by an adjoint Neumann-Ulam implementation while
        the black dashed data was generated by
        Eq~(\ref{eq:analytic_k}). In the right plot, the relative
        difference between the predicted and measured results is
        presented for each weight cutoff.}}
    \label{fig:measured_length}
  \end{spacing}
\end{figure}

@@

To compute the leakage fraction numerically, \sn{3}{5} histories were
sampled from a uniform source of strength unity over the global
domain. At the start of a stage of histories, the number of histories
starting in the center domain was computed and as the stage
progressed, the number of histories that exited that domain was
tallied with the ratio of the two numbers providing a numerical
measure for the leakage fraction. Figure~\ref{fig:measured_leakage}
gives the domain leakage measurements for the domain in the center of
the global grid as well as the analytic result computed by
Eqs~(\ref{eq:wigner_domain_leakage}) and
(\ref{eq:mean_chord_domain_leakage}) as a function of the iteration
matrix spectral radius.
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=6.0in,clip]{chapters/parallel_mc/leakage_variation_2.pdf}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage as a
        function of the iteration matrix spectral radius.} \textit{To
        test the behavior with respect to domain size, $n_i=50$,
        $n_i=100$,and $n_i=200$ were used. The red data was computed
        numerically by a domain-decomposed adjoint Neumann-Ulam
        implementation, the black dashed data was generated by
        Eq~(\ref{eq:mean_chord_domain_leakage}) using the mean-chord
        approximation, and the dashed-dotted black data was generated
        by Eq~(\ref{eq:wigner_domain_leakage}) using the Wigner
        rational approximation.}}
    \label{fig:measured_leakage}
  \end{spacing}
\end{figure}
Again, we note good qualitative agreement between the measured and
analytic quantities but we begin to see the limits of the leakage
approximations. To compare the quality of the two approximations, the
absolute difference between the computed leakage fraction and that
generated by the Wigner rational and mean chord approximations is
plotted in Figure~\ref{fig:leakage_error} for all domain sizes
tested. 
\begin{figure}[t!]
  \begin{spacing}{1.0}
    \begin{center}
      \includegraphics[width=6.0in,clip]{chapters/parallel_mc/leakage_error_2.pdf}
    \end{center}
    \caption{\textbf{Measured and analytic domain leakage absolute
        difference as a function of the iteration matrix spectral radius.}
      \textit{To test the behavior with respect to domain size,
        $n_i=50$ (green), $n_i=100$ (blue), and $n_i=200$ (red) were
        used. The dashed lines represent the difference using the Wigner
        rational approximation while the solid lines represent the
        difference using the mean-chord approximation.}}
    \label{fig:leakage_error}
  \end{spacing}
\end{figure}
From these difference results, the mean chord approximation is shown to
have a lower difference for ill-conditioned systems as compared to the
Wigner approximation while the Wigner approximation produces less
difference for more well-conditioned systems. We also note that for the
optically thick domains, the difference is likely corresponded to that
observed in Figure~\ref{fig:measured_length} for the $k$ parameter
while the large relative difference in $k$ for optically thin domains does
not affect the approximation significantly. In general, the mean chord
approximation is a better choice to estimate the leakage fraction in a
domain from the adjoint Neumann-Ulam method and except for a single
data point with $n_i=50$, the mean chord approximation yielded leakage
fractions within 0.05 of the measured results. As the domain becomes
more optically thick (with both increasing $n_i$ and decreasing
$\rho(\ve{H})$), the approximations are more accurate.

@@

\subsection{Transition Processing Kernel}
\label{subsec:transition_processing}
Processing the transition of a single stochastic history from one
state to another in the system is the fundamental kernel of the
transport procedure. In this kernel, an incoming history has a state
consisting of the current discrete state in the system in which it
resides and its current weight and is processed as given by
Algorithm~\ref{alg:transition_process}.
\begin{algorithm}[h!]
  \caption{Stochastic History Transition Processing Kernel}
  \label{alg:power_iteration}
  \begin{algorithmic}
    \State processTransition( history ):
    \State i = history.state
    \State j = sampleRowCDF( C[i] )
    \State history.state = j
    \State history.weight *= w[i,j]
  \end{algorithmic}
\end{algorithm}
The core component here is the sampling of the CDF for a particular
state generated by the probablities from the Neumann-Ulam
decomposition. Upon exiting, the kernel has modified the state of the
history with a new state and multiplied the weight with the
corresponding transition weight to form the product in
Eq~(\ref{eq:direct_permutation_weight}).
